# Saving SAS Data Sets

```{r, include=FALSE}
library(SASmarkdown)
knitr::opts_chunk$set(SASproctime=FALSE)
```

::: {.callout-tip}
## Saving SAS Data

Save SAS data by either

- using a permanent SAS library
- using a quoted operating system file name
:::

## Introduction
In order for SAS to process any data, that data must be in the format
of a SAS data set.  This is a special SAS binary file which usually
has the file extension ".sas7bdat".  These are typically saved
by writing some SAS code, such as a `DATA` step.

## Temporary and Permanent (Saved) Data Sets
Every SAS data set is stored in your computer's
file system, rather than being solely loaded in your computer's
memory.  However, some data sets are *temporary*
while others are *permanent*. 

Temporary data are written
to a temporary folder usually referred to as the `WORK` library,
which is automatically deleted when SAS is shut down.  Data sets
referred to by a *SAS name* (more below) without a library
reference are stored in the WORK library by default, so they
are temporary.

SAS data sets may be created in several ways, and any of these
may be created as either temporary or permanent.

 - a `DATA` step
 - many `PROC`s have "output data" options
 - all printable output tables can be saved as data using `ODS`


## File References

In your SAS code, these data files can be addressed in
either of two ways, either using an *operating system
name* or using a [*SAS name*](https://documentation.sas.com/doc/en/pgmsascdc/9.4_3.5/lepg/p0ty0ux2rvfhx2n1wka6q16zgtj5.htm).
SAS names are usually preferred, because they are easier to
read, write, and they make your code more portable.

SAS file names generally take the form

```sas
library-name.file-name
```

or just

```sas
file-name
```

The operating system file
name might include the location (the directory or folder)
as well as the file name.  For example a Windows data
file name might look like

```sas
"U:\SAS\class.sas7bdat"
```

while a Linux data file name could look like

```sas
"~/SAS/class.sas7bdat"
```

On either operating system the SAS name for these
data files might look like

```sas
u.class
```

Operating system file names are **ALWAYS** enclosed in
quotes, while SAS file names are **NEVER** in quotes.
In operating systems names, the file path (location) is included
as part of the file name specification.  In SAS file
names, the location is given by a prefix called a
*library* name.  In the SAS example above, `u` is a library
name while `class` is the data file name.

### SAS Names

```{r cleanup, include=FALSE}
win <- (Sys.info()['sysname']=="Windows")
if (win) {
  try(unlink("U:/SAS/class.sas7bdat"))
  } else {
  try(unlink("~/SAS/class.sas7bdat"))
}

```

The basic code to save a SAS data set using the SAS name is

```{sas libwin, eval=win, include=win, collectcode=win}
libname z "U:/SAS";   /* Winstat example */
* libname z "~/SAS";  /* Linstat example */
```
```{sas libunix, eval=!win, include=!win, collectcode=FALSE}
libname z "~/SAS";    /* Linstat example */
* libname z "U:/SAS"; /* Winstat example */
```

```{sas sasname, results="hide"}
data z.class;
  set sashelp.class;
run;
```

This produces a log which says
that the file was successfully saved.

```{saslog sasname, results="hide"}
```

There
are two key elements to this code.  One is the line
that says `data z.class;`, a \"data statement\", which
instructs SAS where to write the file and what to name it.
In this case, the file will be written wherever the
library `z` points to, and will be named `class`.

The other key element, then, is the line that says
`libname z "~/SAS";`, a \"libname statement\", which
tells SAS that the library `z` points to the place
the operating system calls \"~/SAS" (on Linux, the
folder within my home directory).

The data set name and the library name are chosen by
you, but there are certain restrictions (see [SAS names](http://documentation.sas.com/?docsetId=lrcon&docsetTarget=p18cdcs4v5wd2dn1q0x296d3qek6.htm&docsetVersion=9.4&locale=en#p0v4w5zj1972man13kghfmmqec3s)).  SAS names typically begin with a letter
or underscore, and are composed of letters and numerals and
underscores - special characters like blanks or parentheses
are not (usually) allowed.  Library names may be up to
eight characters, while data set names may be up to 32
characters.

(The `set` statement tells SAS where to get data *from*, and 
the `run;` statement tells SAS to execute the instructions
accumulated so far.)

### Operating System Names
Any place you can use a SAS name for a data set, you could
instead use the operating system name.  So the code that
saves the data set, above, could have been written

```{sas osnamewin, eval=win, include=win, results="hide"}
*data "~/SAS/class.sas7bdat"; /* Linstat example */

data "U:/SAS/class.sas7bdat"; /* Winstat example */
  set sashelp.class;
run;
```
```{sas osnameunix, eval=!win, include=!win, results="hide"}
*data "U:/SAS/class.sas7bdat"; /* Winstat example */

data "~/SAS/class.sas7bdat";   /* Linstat example */
  set sashelp.class;
run;
```

And again we'll check the log.

```{saslog osnamewin, eval=win, include=win, results="hide"}
```
```{saslog osnameunix, eval=!win, include=!win, results="hide"}
```

### Pros and Cons
The advantages of using SAS names are

- shorter file references, less clutter
- if you move your project to a new folder or a new platform,
all you have to change is the LIBNAME.

The advantages of using operating system names are

- more flexibile, expressive, file names
- no need for a `libname` statement

In our experience, most SAS programmers use SAS names.